From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-27 upi.tex
not versions
not opt

\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\texttt{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\texttt{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}

@@

@@

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}

{\small
\[\def\arraystretch{1.2}\begin{array}{rcl}
\AgdaFunction{controlled} &:& (a \leftrightarrow a) \rightarrow 
                              \bt \otimes a \leftrightarrow \bt \otimes a \\
\AgdaFunction{controlled}~\AgdaFunction{f} &=& 
      \bt \otimes a \\
&& \qquad\byiso{\AgdaFunction{unfoldBool} \otimes \AgdaFunction{id}} \\
&& (\ot \oplus \ot) \otimes a \\
&& \qquad\byiso{\AgdaFunction{dist}} \\
&& (\ot \otimes a) \oplus (\ot \otimes a) \\
&& \qquad\byiso{\AgdaFunction{id} \oplus (\AgdaFunction{id} \otimes \AgdaFunction{f})} \\
&& (\ot \otimes a) \oplus (\ot \otimes a) \\
&& \qquad\byiso{\AgdaFunction{factor}} \\
&& (\ot \oplus \ot) \otimes a \\
&& \qquad\byiso{\AgdaFunction{foldBool} \otimes \AgdaFunction{id}} \\
&& \bt \otimes a \\
\\ 
\AgdaFunction{not} &:& \bt \leftrightarrow \bt \\
\AgdaFunction{not} &=& 
  \AgdaFunction{unfoldBool} \odot \AgdaFunction{swap₊} \odot \AgdaFunction{foldBool} \\
\\ 
\AgdaFunction{cnot} &:& \bt \otimes \bt \leftrightarrow \bt \otimes \bt \\
\AgdaFunction{cnot} &=& \AgdaFunction{controlled}~\AgdaFunction{not} \\
\\ 
\AgdaFunction{toffoli} &:& \bt \otimes (\bt \otimes \bt)
                           \leftrightarrow  \bt \otimes (\bt \otimes \bt) \\
\AgdaFunction{toffoli} &=& \AgdaFunction{controlled}~\AgdaFunction{cnot}
\end{array}\]}

>>>>

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\byisotwo}[1]{{\Leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\texttt{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\texttt{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}

@@

{\small
\[\def\arraystretch{1.2}\begin{array}{rcl}
\AgdaFunction{controlled} &:& \forall a.~ (a \leftrightarrow a) \rightarrow 
                              (\bt \otimes a \leftrightarrow \bt \otimes a) \\
\AgdaFunction{controlled}~\AgdaFunction{f} &=& 
      \bt \otimes a \\
&& \qquad\byiso{\AgdaFunction{unfoldBool} \otimes \AgdaFunction{id}} \\
&& (\ot \oplus \ot) \otimes a \\
&& \qquad\byiso{\AgdaFunction{dist}} \\
&& (\ot \otimes a) \oplus (\ot \otimes a) \\
&& \qquad\byiso{\AgdaFunction{id} \oplus (\AgdaFunction{id} \otimes \AgdaFunction{f})} \\
&& (\ot \otimes a) \oplus (\ot \otimes a) \\
&& \qquad\byiso{\AgdaFunction{factor}} \\
&& (\ot \oplus \ot) \otimes a \\
&& \qquad\byiso{\AgdaFunction{foldBool} \otimes \AgdaFunction{id}} \\
&& \bt \otimes a \\
\\ 
\AgdaFunction{not}~\AgdaFunction{not₁}~\AgdaFunction{not₂}~\AgdaFunction{not₃} &:& 
     \bt \leftrightarrow \bt \\
\AgdaFunction{not} &=& 
  \AgdaFunction{unfoldBool} \odot \AgdaFunction{swap₊} \odot \AgdaFunction{foldBool} \\
\AgdaFunction{not₁} &=& 
  \AgdaFunction{id} \odot \AgdaFunction{not} \\
\AgdaFunction{not₂} &=& 
  \AgdaFunction{not} \odot \AgdaFunction{not} \odot \AgdaFunction{not} \\
\AgdaFunction{not₃} &=& 
  \AgdaFunction{uniti⋆} \odot \AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{not} \otimes \AgdaFunction{id}) \odot 
                        \AgdaFunction{swap⋆} \odot 
                        \AgdaFunction{unite⋆} \\
\\ 
\AgdaFunction{not} &:& \bt \leftrightarrow \bt \\
\AgdaFunction{not} &=& 
  \AgdaFunction{unfoldBool} \odot \AgdaFunction{swap₊} \odot \AgdaFunction{foldBool} \\
\\ 
\AgdaFunction{cnot} &:& \bt \otimes \bt \leftrightarrow \bt \otimes \bt \\
\AgdaFunction{cnot} &=& \AgdaFunction{controlled}~\AgdaFunction{not} \\
\\ 
\AgdaFunction{toffoli} &:& \bt \otimes (\bt \otimes \bt)
                           \leftrightarrow  \bt \otimes (\bt \otimes \bt) \\
\AgdaFunction{toffoli} &=& \AgdaFunction{controlled}~\AgdaFunction{cnot} \\
\\
\AgdaFunction{notOpt} &:& \AgdaFunction{not₃} \Leftrightarrow \AgdaFunction{not} \\
\AgdaFunction{notOpt} &=& 
  \AgdaFunction{uniti⋆} \odot (\AgdaFunction{swap⋆} \odot 
                        ((\AgdaFunction{not} \otimes \AgdaFunction{id}) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆}))) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{assocLeft}} \\
&& \AgdaFunction{uniti⋆} \odot (\AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{not} \otimes \AgdaFunction{id})) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆}) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{swapLeft} 
                                  \boxdot \AgdaFunction{id})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot \AgdaFunction{swap⋆}) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆}) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{assocRight}} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot (\AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆}))) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot \AgdaFunction{assocLeft})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot ((\AgdaFunction{swap⋆} \odot 
                      \AgdaFunction{swap⋆}) \odot \AgdaFunction{unite⋆})) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot (\AgdaFunction{leftInv} \boxdot \AgdaFunction{id}))} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot (\AgdaFunction{id} \odot \AgdaFunction{unite⋆})) \\
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot \AgdaFunction{idLeft})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot \AgdaFunction{unite⋆}) \\
&& \qquad\byisotwo{\AgdaFunction{assocLeft}} \\
&& (\AgdaFunction{uniti⋆} \odot (\AgdaFunction{id} \otimes \AgdaFunction{not})) 
                      \odot \AgdaFunction{unite⋆} \\
&& \qquad\byisotwo{\AgdaFunction{unitiLeft} \boxdot \AgdaFunction{id}} \\
&& (\AgdaFunction{not} \otimes \AgdaFunction{uniti⋆}) \odot \AgdaFunction{unite⋆} \\    
&& \qquad\byisotwo{\AgdaFunction{assocRight}} \\
&& \AgdaFunction{not} \otimes (\AgdaFunction{uniti⋆} \odot \AgdaFunction{unite⋆}) \\    
&& \qquad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{leftInv}} \\
&& \AgdaFunction{not} \otimes \AgdaFunction{id} \\
&& \qquad\byisotwo{\AgdaFunction{idRight}} \\
&& \AgdaFunction{not} 
\end{array}\]}

@@
